---
title: "Helicpoter Final Experiment"
output: html_document
date: "2024-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Introduction 



## Experimental Design 

The five variables we chose to experiment with are wing length, body length, body width, drop style, and whether or not we drop with a paperclip. The levels of these factors are as follows

* A = wing_length: [4,5]
* B = body_length: [5,6]
* C = body_width: [3.5,5]
* D = drop_style: [pinched wings, open]
* E = paper_clip: [paper clip, no paper clip]


Our experiment is a half factorial design created using the code below. The design is created with 16 helicopters and replicated once for 32 helicopters. The -1 refers to the first level of each factor and the 1 refers to the second level of the factor. 


```{r}
library(FrF2)
d<-FrF2(16, 5, randomize=FALSE)
d$y<-rnorm(16)
aliases( lm( y~ (.)^4, data = d))
```
The aliases show that we have a resolution IV design, which is ideal for this experiment


```{r}
d<-FrF2(16, 5, randomize=FALSE)
d<-as.data.frame(d)
d<-rbind(d, d)
d
```
## Randomizing Run Order

```{r}
set.seed(633)  # Setting seed for reproducibility
shuffled_experiments <- d[sample(nrow(d)), ]
shuffled_experiments
```

## Experiment


