---
title: "Homework 3"
date:  "Last complied on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
  
    
    theme: sandstone
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

1.  Refer to the Kohavi et al (2009) article. Look at the example in Section 3.2.1 on page 153. Verify the sample size calculations using the Z statistics for $\alpha/2$=0.025 and $\beta$=0.2 and this formula: $$n_1=\frac{2(Z_{(1-\alpha/2)}+Z_{(1-\beta)} )^2\times \sigma^2}{\Delta^2} $$

They give an approximation in their example. Compute for both responses, userspend and conversion rate. Explain the difference between these sample size numbers.

```{r}
# Users to achieve 80% power with 5% change in user spend
((2*(1.96 + .7881)^2) * 30^2)/(3.75 * 0.05)^2
```

Now looking at a 5% change in conversion rate:

```{R}
# Users to achieve 80% power with 5% change in conversion rate
((2*(1.96 + .7881)^2) * (0.05*(1-0.05)))/(0.05*0.05)^2
```

The difference in the sample size numbers comes from the use of the Z scores in our equation. The book uses a simpler formula using 16 \* the standard deviation divided by the difference squared. Our formula calculates the required sample sizes difference which explains the slights differences.

Investigate the following relationships in terms of statistical power to detect a difference between two means. For example, below is a plot that illustrates how changing effect size (mean difference) changes the power. Using this framework, investigate the following relationships. Use power=0.8, sd=30, n=100, and $\alpha$=0.05 unless otherwise stated.

```{r}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=100, sd=30, sig.level=0.05, power=NULL)$power
}

plot(d, results, xlab="delta", ylab="power")
abline(h=0.8, col="red")


```

2.  What happens to the power when the standard deviation increases and nothing else changes?

```{r}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=100, sd=40, sig.level=0.05, power=NULL)$power
}

plot(d, results, xlab="delta", ylab="power")
abline(h=0.8, col="red")


```
Power decreases

3.  What happens to the power when we increase sample size and nothing else changes?

```{r}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=200, sd=30, sig.level=0.05, power=NULL)$power
}

plot(d, results, xlab="delta", ylab="power")
abline(h=0.8, col="red")


```
Power increases

4.  What happens to the power when change the $\alpha$ level and nothing else changes?

```{R}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=100, sd=30, sig.level=0.1, power=NULL)$power
}

plot(d, results, xlab="delta", ylab="power")
abline(h=0.8, col="red")

```
When the $\alpha$ level goes up, power increases

```{r}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=100, sd=30, sig.level=0.025, power=NULL)$power
}

plot(d, results, xlab="delta", ylab="power")
abline(h=0.8, col="red")
```

When the $\alpha$ level decreases, power decreases

5.  What happens to the sample size when we hold power=0.8 and increase the effect size?

```{R}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=NULL, sd=30, sig.level=0.05, power=0.8)$n
}

plot(d, results, xlab="delta", ylab="sample size")

```
```{r}
results<-c()
d<-c(5, 10, 15, 30, 35, 40, 45, 50, 60, 65)
for(i in 1:length(d)){
  results[i]<-power.t.test(delta=d[i], n=NULL, sd=30, sig.level=0.1, power=0.8)$n
}

plot(d, results, xlab="delta", ylab="sample size")
```
Sample size decreases

You are on a marketing team at Miami University working on a project to increase brand recognition. You suggest running an email experiment where a random sample of prospective students gets one email version and another sample receives a new version. The current rate that students respond to the email is 0.23%. The Miami University marketing group would like to increase this rate by a minimum of 5%.

6.  How many emails will you have to send to each group to achieve this percentage increase?

```{r}
power.prop.test(n=NULL, p1=0.23, p2=0.28, sig.level = 0.05, power=.8)
```
You would have to send at least 1,192 emails to each group to achieve this percentage increase

7.  Is this a feasible increase to target? Why or why not?

Yes, as this is using a feasible significane level and power, and sending 2,384 emails to a large university is not unfeasible

In a hypothetical case, a promotion for a hair care produce was sent to some members of a buyers club. Purchases were then recorded for both the members who got the promotion and those who did not. Use the data "Hair-Care-Product.csv" to answer the following questions.

```{r}
df<-read.csv("https://raw.githubusercontent.com/losiewbw/ISA633/main/airbnb_social.csv")
```


8.  What is the purchase propensity for those who did and did not received the promotion?

```{r}




```

9.  Partition the data into training and validation using the Validation column. Fit an uplift model using a random forest. Do this in Python using the `reticulate` package or copy these questions to a Jupiter notebook and turn in another separate html file.

10. Find the top 10 persuadables in the data set describe these people.
